### 2021-10-07 01_data/newsarticle/counting-countries2.Rmd
r codes for counting countries mentioned
Now we can do separate years! Oh even more convenient!

---
title: "Preprocessing: counting countries Xinhua 2000"
author: "Lucie Lu"
date: "10/6/2021"
output: html_document
---

```{r}
test <- read.csv("Xinhua_2000.csv")
#head(test)
library(tidyverse)
#https://hbctraining.github.io/Intro-to-R/lessons/08_intro_tidyverse.html
test <- test %>% 
  rownames_to_column() %>% 
  as_tibble()
```

```{r}
df <- test %>% 
  pull(	
country) %>% 
  str_split("[|]")

max(lengths(df))

names(df) <- test$rowname
#df_2 <- data.frame(matrix(unlist(df), nrow=max(lengths(df)), byrow=TRUE),stringsAsFactors=FALSE)

library (plyr)
df_2 <- ldply (df, data.frame)
df_2 <- tibble(df_2)
colnames(df_2)[2] <- "country_mentioned"

df_2$countries_mentioned <- as.factor(df_2$country_mentioned)
```

```{r}
library(dplyr)
df_2 %>% 
  group_by(country_mentioned) %>%
  dplyr::summarise(n = n()) %>%
  arrange(desc(n))

#duplicated name summarise in both packages
#https://statisticsglobe.com/r-error-must-only-be-used-inside-dplyr-verbs
```