### 2021-12-15 03_analysis/02_results_summary.Rmd
some preliminary results

---
title: "Results"
#author: "Miles D. Williams"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

**Note**: Table 1 shows results with the imputed dataset. Table 2 shows results with the non-imputed dataset.

```{r, results='asis'}
load(
  paste0(getwd(), "/tab1.R")
)
reg.tab1 <- reg.tab
cat(reg.tab1)
```

```{r, results='asis'}
load(
  paste0(getwd(), "/tab2.R")
)
reg.tab2 <- reg.tab
cat(reg.tab2)
```