### 2021-12-19 03_analysis/ElasticNet.Rmd
Add files via upload

---
title: "ElasticNet and Lasso"
author: "Lucie Lu"
date: "12/18/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

data <- read.csv("https://raw.githubusercontent.com/milesdwilliams15/chinese_media_and_foreign_aid/main/01_data/final_data/final_data_imputed.csv")

library(Matrix)

set.seed(2234653)
split_data <- sample(nrow(data), floor(0.7*nrow(data)))
train_data <- data [split_data,]
test_data <- data [-split_data,]

library(glmnet) 
#train_sparse_3_sub <- sparse.model.matrix(~.,train_3_sub)
#test_sparse_3_sub <- sparse.model.matrix(~.,test_3_sub)

library(caret)

#set training control
train_control <- trainControl(method="cv", number = 10, 
                              search="random",
                              verboseIter=T)

sel_x <- train_data[c("count", "pmm_polity", "gdp", "pop", "unemp", "disaster", "civilwar",
                          "dist", "trade", "atopally")]

# tune the variables to get lambda and alpha for Elastic Net
fit_en <- train(y=train_data$aid, x=as.matrix(sel_x),
                              method="glmnet", tuneLength=25, trControl=train_control)

fit_elasticnet_f <- glmnet(y=train_data$aid, x=as.matrix(sel_x), 
                                 lambda = 2.17, alpha = 0.031)

coef(fit_elasticnet_f)

```

```{r}

sel_x_t <- test_data[c("count", "pmm_polity", "gdp", "pop", "unemp", "disaster", "civilwar",
                          "dist", "trade", "atopally")]

pred_elasticnet_f <- predict(fit_elasticnet_f, as.matrix(sel_x_t))
rmse_f <-  sqrt(mean((pred_elasticnet_f - test_data$aid)^2))
rmse_f

#this is very huge.....
library(ModelMetrics)
rmse(test_data$aid, pred_elasticnet_f)
```

```{r ridge}

set.seed(323445)
fit2 <- cv.glmnet(x = data.matrix(sel_x), y = train_data$aid, nfolds = 10, alpha = 1)
plot(fit2$glmnet.fit, "lambda")
plot(fit2) #MSE is huge here

# using the lambda minimizing the cross-validation error leaves us only the "count" variable. Odd.
coef(fit2, s = "lambda.min")

# nothing is left here... This gives a larger penalty with more shrinkage. 
coef(fit2, s = "lambda.1se")

```