### 2022-04-06 05_Presentation/MPSA_AID_Paper_Presentation.tex
Add files via upload

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathabx}
\usepackage{mathpazo}
\usepackage{eulervm}
\usepackage{natbib}

%% Load the markdown package
\usepackage[citations,footnotes,definitionLists,hashEnumerators,smartEllipses,tightLists=false,pipeTables,tableCaptions,hybrid]{markdown}
%%begin novalidate
\markdownSetup{rendererPrototypes={
 link = {\href{#2}{#1}},
 headingOne = {\section{#1}},
 headingTwo = {\subsection{#1}},
 headingThree = {\begin{frame}\frametitle{#1}},
 headingFour = {\begin{block}{#1}},
 horizontalRule = {\end{block}}
}}

%%end novalidate

\usetheme{Dresden}
\usefonttheme{serif}
\usecolortheme{rose}

\title{Friends Get More Money, Attention and Handshakes:
Chinese Foreign Aid, Xinhua and Diplomatic Visits}

\author{Lucie Lu, Miles D. Williams\\ University of Illinois at Urbana-Champaign}
\date {MPSA 2022 \\ April 8, 2022}

\begin{document}

\maketitle

\begin{frame}{Puzzle}

\begin{itemize}
    \item China seems to deliberately keep its foreign aids secret
    \item However, aids as part of the "South-South cooperation" improve China's image abroad.
    \item Two opposite modes of interactions: foreign aids and media/public diplomacy -- secret and open
\end{itemize}

\end{frame}

\begin{frame}{Research Question and Argument}

\begin{itemize}
    \item Research Question: How does China’s foreign aid relate to Beijing’s efforts at public diplomacy and international recognition?

   \item Argument: Countries that receive higher media exposure and diplomacy efforts are
strategically more important to Beijing, and in turn receive disproportionally larger China’s
foreign aid.

\end{itemize}
\end{frame}

\begin{frame}{Argument Continued}

Two dimensions of Chinese public diplomacy efforts:

\begin{itemize}
    \item state-sponsored \textit{Xinhua} mediac coverage of developing countries intended for foreign audiences
    \item China’s bilateral South-South diplomatic activities

\end{itemize}

Developing countries that receive \textbf{higher} media coverage and diplomatic visit are strategically more important than the others. Hence, they get \textbf{more} benefits of aids.

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Motivations behind Chinese Foreign Aid}

Mirror Western donors:

\begin{itemize}
    \item Countries in need \begin{footnotesize} (Dreher et al. 2018) \end{footnotesize}
    \item Countries import more goods from China \begin{footnotesize} (Dreher and Fuchs, 2015) \end{footnotesize}
    \item Countries with more natural resources, closer voting alignment with Beijing, more capacity to repay loans \begin{footnotesize} (Dreher et al., 2021) \end{footnotesize}
\end{itemize}

Similar to Western donors, the transparency of foreign aid arguably serve China’s interests.
\end{frame}

\begin{frame}{Motivations behind Chinese Foreign Aid}
The obscurity seems to have more to do with capacity and logistics, rather than intention \begin{footnotesize} (Dreher et al. 2018) \end{footnotesize}. We
explore the \textbf{connections }between China’s bilateral development finance and other, visible,
activities linked to diplomatic and legitimacy-seeking objectives.

\end{frame}

\begin{frame}{Media: First-level Agenda-setting}

\begin{itemize}
\item What's covered in the news signal to the public what is important \begin{footnotesize} (McCombs \& Shaw, 1972) \end{footnotesize}

\item Media coverage of the countries will lead to the countries as a whole becoming more
salient among the public. \begin{footnotesize} (Wanta et al., 2004) \end{footnotesize}

\item News editors as gatekeepers: the increased media salience suggests the salience of the foreign countries to the audiences and the \textbf{newsworthiness }to the news editors.
\end{itemize}

\end{frame}

\begin{frame}{Media: Second-level Agenda-setting}

\begin{itemize}
\item Media are critical in building states’ image to international audiences \begin{footnotesize} (Nye, 2008) \end{footnotesize}

\item Media coverage shapes public attitudes -- core objective of mediated public diplomacy. \begin{footnotesize} (Entman, 2008; Kiousis \& Wu, 2008) \end{footnotesize}

\item Xinhua has dual roles: reporting news and building the China
image as part of Beijing’s “going out” strategy \begin{footnotesize} (Shambaugh, 2015) \end{footnotesize}

\item Its primary function is to transmit information. It reports on a variety of global issues
with different regional focuses to meet the standard of international news competitors -- \textbf{editors' choices} of what countries get more coverage
\end{itemize}

\end{frame}

\begin{frame}{Public Diplomacy}

\begin{itemize}
\item Visiting leaders can increase the awareness of themselves and their country among
citizens in the host country \begin{footnotesize} (Goldsmith \& Horiuchi, 2009; Goldsmith et al., 2021) \end{footnotesize}

\item Diplomatic visits usually come with policy agenda between two countries,
and the two are expected to cooperate in various domains through their negotiations.

\item Diplomatic visit is a signal of support and proximity of
the countries.

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Hypotheses}

\begin{itemize}

\item Both media exposure and public diplomacy involve limited time and resources,
we expect that Beijing will choose on whom to cast the spotlight, and whom to visit,
strategically.

\item We expect China to prioritize giving foreign
aid to its “friends” -- countries where China has developed strong relationships through
public diplomacy efforts.

\end{itemize}

\end{frame}

\begin{frame}{Hypotheses}

\begin{itemize} 
    \item \textit{H1:} When developing countries receive more media exposure in Xinhua, they are more
likely to receive more foreign aid from China.
    \item \textit{H2:} When developing countries host or visit Chinese high-level diplomatic visits more
frequently, they are more likely to receive more foreign aid from China.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Data and Methods}

Unique recipient-year observations from 2000 to 2014

\begin{itemize}

\item DV: bilateral aid data (AidData’s Chinese development finance dataset)

\item IV: number of diplomatic visits (AidData’s compiled yearly counts of bilateral official
diplomatic visits from China)

\item IV: counts of country media coverage (Millions of English edition Xinhua news articles scraped by the Cline Center for Advanced Social Research)

\end{itemize}

\end{frame}

\begin{frame}{Data and Methods}

\begin{itemize}

\item Covariates: aid recipient characteristics (GDP, presence of war, disasters, political and civil liberties) and bilateral characteristics between China and aid recipients (economic ties, alliances)

\item Model Specifications: OLS with robust standard errors clustered by recipients; Tobit maximum likelihood estimator; selection and level equations; instrumental variables

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{Results}

\end{frame}

\end{document}