### 2022-06-12 04_paper/Draft2.Rmd
more changes to draft 2

---
title: "Friends Get More Money, Attention and Handshakes: Chinese Foreign Aid, Xinhua and Diplomatic Visits"
author: ["Lucie Lu","Miles D. Williams"]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    number_sections: yes
classoption: fleqn
fontfamily: mathpazo
fontsize: 12pt
bibliography: ref_Xinhua.bib
link-citations: yes
linkcolor: blue
header-includes:
  - \usepackage{amsmath}
  - \usepackage{setspace}
  - \setlength{\mathindent}{15pt}
  - \setlength\parindent{24pt}
  - \setlength{\parskip}{0.0pt plus 1.0pt}
  - \doublespacing
  - \newtheorem{theorem}{Theorem}[section]
  - \newtheorem{lemma}[theorem]{Lemma}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \newtheorem{corollary}[theorem]{Corollary}
  - \newenvironment{proof}[1][Proof]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries Proof}]}{\end{trivlist}}
  - \newenvironment{definition}[1][Definition]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{example}[1][Example]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newenvironment{remark}[1][Remark]{\begin{trivlist}
      \item[\hskip \labelsep {\bfseries 1}]}{\end{trivlist}}
  - \newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.75em width0.5em depth0.25em\fi}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \floatplacement{figure}{!t}
  - \usepackage{caption}\captionsetup[table]{skip=5pt}
  - \usepackage{dcolumn}
  - \usepackage{rotating}
  - \usepackage{caption}\DeclareCaptionLabelFormat{AppendixTables}{A.#2}

abstract: "China's international development finance has been long suspecpted to serve Beijing's efforts to promote its legitamcy and image on the world stage and to support closer ties with aid recipients. But what is puzzling is that these goals require transparency and publicity of development projects, which are two virtues China's development finance has historically lacked. We bring a novel perspective to this issue by considering how China's foreign aid complements more visible foreign policy levers Beijing may use to accomplish its goals: (1) state run media coverage of developing countries intended for foreign audiences and (2) bilateral diplomatic visits. These activities capture two distinct but related aspects of China’s foreign policy: status/legitimacy in the eyes of a foreign (predominantly Western) audience and south-south diplomacy. To the extent that China’s development finance is intended to support these goals, we expect greater coverage of a developing country in Chinese media directed to foreign readers and greater bilateral diplomatic activity to correlate with greater Chinese foreign aid. To test this expectation we merge AidData’s Chinese development finance dataset with two newly available datasets: (1) AidData’s compiled yearly counts of bilateral official diplomatic visits from China and (2) millions of English edition *Xinhua* news articles from 2000 to 2014 scraped by the Cline Center for Advanced Social Research. We find that greater media coverage of developing countries targeted at a foreign audience via the English version of *Xinhua* and the number of diplomatic visits to a developing country predict greater receipt of Chinese aid. Our results support the view that Chinese aid allocation maps to China’s broader diplomacy and legitimacy seeking objectives on the world stage."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

# Introduction

China neither reports its foreign aid spending to international organizations, as do Development Assistance Committee (DAC) members to the OECD, nor does China publish its aid expenditures in a public database, as does the United States in USAID's *Greenbook*. @fuchsRudyak2019 cite this fact as a reason Western governments assume Beijing deliberately keeps its international development finance a secret.

While limited transparency is the norm for Beijing, in the context of its global development finance the lack of transparency is a puzzle. Many observe an unmistakable connection between China's version of Official Development Assistance (ODA) and its political and material foreign policy goals. These ambitions include international recognition and closer diplomatic ties with recipients in what China calls "South-South cooperation." While distinct, these objectives have one thing in common: they are difficult to accomplish *in secret.* They require visibility. So, absent traditional modes of transparency, *how does China's foreign aid relate to Beijing's efforts at public diplomacy and international recognition?*

We probe this question empirically by exploiting two newly available datasets of Chinese media coverage of developing countries intended for foreign audiences and of China's bilateral diplomatic visits, respectively. These activities capture two distinct but related aspects of China's foreign policy: status/legitimacy in the eyes of a foreign (predominantly Western) audience and south-south diplomacy. To the extent that China's development finance is targeted at these goals, we expect China to complement greater state-run media coverage of developing countries and greater bilateral diplomatic activity with greater development finance.

To test this expectation, we merge AidData's Chinese development finance dataset with: (1) AidData's compiled yearly counts of bilateral official diplomatic visits from China and (2) millions of English edition *Xinhua* news articles from 2000 to 2014 scraped by the Cline Center for Advanced Social Research. We find that greater media coverage of developing countries targeted at a foreign audience via the English version of *Xinhua* and the number of diplomatic visits to a developing country predict greater receipt of Chinese aid. Our results support the view that Chinese aid allocation patterns map to China's broader diplomacy and legitimacy-seeking objectives on the world stage.

This work further helps in dispelling the notion that China purposely engages in development finance in secret. Much to the contrary, China's development finance moves in step with visible acts of diplomacy and external-facing media coverage. To our understanding, this is the first attempt to use media agenda-setting and public diplomacy to explain Chinese foreign aid allocations. As we show, government-sponsored communication, diplomatic outreach, and financial assistance to developing countries are all crucial to Beijing achieving its foreign policy goals and promoting its soft power.

# Motivations behind Chinese Foreign Aid

China's motivations for giving foreign aid to developing countries mirror, in many ways, the motivations of Western donors. @bermeo2018 provides a recent and authoritative review of the literature on donor objectives, which have historically centered on some combination both of non-development geopolitical and material interests, and of legitimate efforts to address recipient country needs. Beijing's aid giving appears to follow a similarly mixed set of motivations. Studies show that China targets more finance toward poorer countries and does not systematically favor authoritarian over democratic recipients [@dreherEtAl2018]. At the same time, China gives more aid to countries that import more goods from China, while it gives less aid to those that officially recognize Taiwan [@dreherFuchs2015]. Further, China's Official Financing disproportionately goes to countries with more natural resources, greater UN General Assembly voting alignment with Beijing, and with more capacity to repay loans [@dreherEtAl2021].

Two waves of surveys conducted by AidData of policymakers in 126 countries demonstrate the diplomatic impact that these investments have had.[^1] Compared to a first survey wave in 2014, responses in 2017 showed an increase in evaluations of China's influence with policymakers in developing countries. It would be naive to suppose this is a side-effect, rather than a goal, of China's development finance and other investments.

[^1]: Reported by AidData ["China's Financial Statecraft: Winning Africa one Yuan at a Time?"](https://www.aiddata.org/blog/chinas-financial-statecraft-winning-africa-one-yuan-at-a-time) on March 22, 2018. Accessed March 3, 2022

If its diplomatic motives for giving aid parallel those of other donor governments, visibility is surely a major concern for Beijing. Visual branding of aid has emerged as a growing concern among donors. The United States Agency for International Development (USAID), for example, cited its "brand identity" as instrumental in promoting the visibility of 2004-2005 tsunami relief to Indonesia.[^2] A US State Department report from 2004 further claims that better branding contributed to a doubling of favorable attitudes toward the US. Pew Research supports this view, finding that nearly 80$\%$ of Indonesians surveyed "said that post-tsunami aid from the US had improved their impression of America."[^3]

[^2]: See USAID's resource page on branding: [usaid.gov/branding](https://www.usaid.gov/branding).

[^3]: ["Does humanitarian aid improve America's image?"](https://www.pewresearch.org/global/2012/03/06/does-humanitarian-aid-improve-americas-image/) Published by Pew Research March 6, 2012. Accessed March 3, 2022.

Given the diplomatic uses of foreign aid, the obscurity that surrounds Beijing's development finance is all the more puzzling. Unlike Western donors, China does not report its financing activities to an international organization or make its activities visible in an online database. To say transparency is out of character for Beijing is an understatement, but foreign aid is one issue where transparency would arguably serve China's interests. So why the secrecy?

As @fuchsRudyak2019 point out, the obscurity of China's development finance creates the perception among Western donors that China purposefully keeps its aid giving a secret. But as several scholars note, lack of transparency may have more to do with capacity and logistics than intention.[^4] For many decades, a complex and fractured bureaucracy has orchestrated Beijing's development financing. Only in the past few years, as its goals have become ever loftier, did China establish its first true bilateral aid agency---the China International Development Cooperation Agency (CIDCA). While it was established in 2018, much work remains to consolidate the operation and management of China's bilateral aid.[^5]

[^4]: See @fuchsRudyak2019 for several related citations.

[^5]: See this excellent summary entitled ["The Logic Behind China's Foreign Aid Agency"](https://carnegieendowment.org/2019/05/21/logic-behind-china-s-foreign-aid-agency-pub-79154) by the Carnegie Endowment for International Peace published May 21, 2019. Accessed March 10, 2022.

Absent transparency, how does Beijing leverage its global development finance in support of its diplomatic interests and international recognition? In this paper, we seek out an answer by looking at other kinds of visible foreign policy activities. We exploit two newly available datasets to explore the connections between China's bilateral development finance and critical *visible* activities linked to the diplomatic and legitimacy-seeking objectives scholars attribute to China's international development finance. We focus in particular on two factors: (1) diplomatic visits from China to developing countries and (2) coverage of developing countries in external-facing state-run media.

The details of these factors are covered in greater detail in the following sections. However, to quickly summarize our argument, we believe diplomatic visits and media coverage are important factors for two reasons.

First, they signal the importance of developing countries to Beijing. Just like development finance, diplomatic visits and state-run media coverage are subject to scarcity---the first in terms of limited resources and bandwidth; the second in terms of time dedicated to writing, editing, and publishing content. Within the rationalist framework most often adopted by IR scholars who study international aid,[^6] this implies a direct correspondence between the countries that host diplomatic visits and that are subjects of state-sponsored news articles and the importance of developing countries to China. In short, visits and coverage are costly, and they therefore tell us something about which developing countries Beijing values most.

[^6]: @bermeo2018 and @BDM2009 are exemplary examples of IR research in international aid that falls within the rationalist tradition.

Second, diplomatic visits and state-sponsored media coverage are *visible* acts. Though some diplomatic visits are certainly covert, most are public, even performative displays of the closeness and alignment between host and guest. And media coverage are by their very nature public.

Thus, taken together, diplomatic visits to, and state-run media coverage of, developing countries serve both as signals of the importance of developing countries to Beijing and as visible indicators of the diplomatic and legitimacy-seeking objectives proposed to motivate China's international development finance activities.

# Xinhua and Setting the Agenda

Media agenda-setting is critical in building a state's image to international audiences. @nye_public_2008-2 suggests that the first and the most immediate dimension of investing in public diplomacy is through daily communications with the aim to improve a country's "soft power". Government officials devote a great deal of attention to what and how to tell the press, both the domestic press and the foreign press. For this reason, China has expanded its *Xinhua* News Agency to reach global audiences during the past decade to take greater control of how it is portrayed by the international press.

As a state-sponsored media platform, *Xinhua* News Agency is often considered more than a news media agency. It has had a dual role: reporting news and building up China's image abroad as part of Beijing's "going out" strategy [@shambaugh_chinas_2015-1]. It has opened over 170 bureaus across the globe and competes head-to-head with the main Western newswires. Among many audiences, the agency targets developing countries, where the Western press has contracted out in recent years due to the decline of advertisement revenues. It is an international news agency that tries to alter its stilted and propagandistic flavor and package its content in a more viewer-friendly format. Beijing uses media to tell its story to the world and improve its public relations with other countries. As such, scholars often focus on Chinese mediated public diplomacy efforts through *Xinhua*. That is, they study how government-sponsored international broadcasting contributes to influencing public or elite opinion among foreign countries [@entman_theorizing_2008; @manheim_strategic_1994; @sheafer_mediated_2009)]

Despite growing in the global media landscape, *Xinhua*, as an international broadcaster, needs foreign audiences. Simply providing more information will not improve their image in the Western world. As @druckman_limits_2001 notes, one conditional factor in influencing public opinion (especially foreign audiences'), is an information sender's "credibility." On the one hand, governments attempt to positively shape the global narrative regarding their leaders and foreign policies using their own newsgathering and broadcasting operations. On the other hand, *Xinhua* needs to maintain low political bias and higher professional standards to assert its brand and increase its international competitiveness.

Hence, while *Xinhua* plays a major role in defining, shaping, and projecting China's image, its primary function is to transmit information. It reports on a variety of global issues with different regional focuses to meet the standard of international news competitors.

However, mere information transmission is rarely a neutral act. It reflects certain priorities, which in turn effect the priorities and perceptions of audiences. Research on the first-level and second-level agenda-setting power of media illustrates this fact.

## First and Second Level Agenda-Setting

Agenda-setting refers to the ability of media to signal to the public what is important. Media coverage of a foreign country measures how important a country is to its reporting country and, in particular, to the public. The public learns the importance of issues based on the amount of coverage that those issues receive. Agenda-setting theory suggests that mass media serve as one of the key sources for public perception of important issues [@mccombs_agenda-setting_1972]. More specifically, as @berelson_mass_1966 observe, "The mass media force attention to certain issues. They build up public images of political figures. They are constantly presenting objects suggesting what individuals in the mass should think about, know about, have feelings about" (p. 468). Since their seminal work, hundreds of studies have examined this media effect on the public. The vast majority support this agenda-setting notion (need some more extension or citations here).

Over decades of rich research in agenda-setting, empirical patterns support Cohen's [-@cohen_press_1963] idea that the media is "stunningly successful in telling its readers what to think *about*" (p. 13; emphasis added). That is, the salience of news media reports transfers to the perception of issue salience to the public [@mccombs_agenda-setting_1972]. Extensive media coverage provides news consumers with salience cues regarding the importance of the political figures or issues, and in turn shape individuals' perceptions of the relative importance and salience of the issues.

To extend this logic, media salience indicates the significance of foreign countries named in the media. @wanta_agenda_2004-3 show that coverage of foreign nations in the news in the US relates to respondents' perception of the importance of the countries: the more coverage a state receives, the more likely respondents think the state is vitally important to the US. The media coverage of the countries will lead to the countries as a whole becoming more salient among the public.

Recent research on what is called second-level agenda-setting suggests that news media can influence how people think about a topic by emphasizing certain issues or subjects over and above others [@entman_theorizing_2008; @kiousis_international_2008]. Compared to the first-level agenda-setting, this mechanism entails a deeper media influence on shaping public attitudes. Rather simply focusing on what the public thinks about more, second-level agenda setting emphasizes media's ability to shape how the public think about issues. In the case of *Xinhua*, this implies an ability to shape how foreign audiences perceive the countries covered by the news service.

## Editorial Gatekeeping and Scarcity

When it comes to agenda-setting, the *editor* is a critical source of gatekeeping power. From the perspective of news outlets, limited by time and space, editors can only select a handful of stories while relegating many others to file drawer. News selection is at the heart of the agenda-setting process since the issues that fail to pass through the gatekeepers of the news also fail to give salience cues regarding the relative importance of the issues.

Also, previous findings show the strategic consideration of news editors in considering what international news gets media coveriage. Not all countries in the world receive equal coverage. In fact, only a small portion of international events get through the media gatekeepers. While powerful core states consistently receive higher amounts of coverage from newscasts, small peripheral states remain largely uncovered. @golan_international_2003-1's study examines how 138 elections held between 1998 and 2000 were covered by US network television newscasts (ABC, CBS, NBC and CNN). They found that elections in Europe, Asia or the Middle East received more substantial coverage than those in Latin America and Africa. A large body of literature has examined the determinants of international news coverage and has identified variables such as geographic locations, cultural proximity, and geopolitical position in the world system as key predictors of coverage [@sheafer_relative_2014-1]

It seems unlikely that *Xinhua* would be an exception to the rule that biases exist in the gatekeeping decisions of editors. However, in the case of *Xinhua*, given the discussion of its role in China's "going out" strategy, it seems reasonable that its bias will be in favor of countries that are important to Beijing. In particular, it will cover countries that are important to Beijing's foreign policy goals.

# Public Diplomacy

The previous section illustrates that media agenda-setting is an essential step in a country's public diplomacy as it tries to influence the perceptions of a foreign audience. To borrow from one media scholar, external facing media allow Beijing to "speak to [foreign] people -- and listen to them" [@hoffman_introduction_1968, 4] In fact, all public diplomacy programs have a primary responsibility to explain and defend government policies to foreign audiences" [@deibel_culture_1976, 15].

A traditional mental image of public diplomacy is "diplomats engaged in traditional negotiations under the glare of publicity" [@deibel_culture_1976, 13]. In addition to negotiations behind closed doors, records of visits provide symbolic evidence of the bilateral relationship between countries. Perhaps most importantly, government-to-government visits are not costless---they require time and resources to orchestrate and enact. They often are meant to develop and strengthen bilateral relations, especially with respect to bilateral trade [@nitsch_state_2005]. Leader visits also have implications on the host leaders' security in office [@malis_state_2021] and on the host state's security in general [@mcmanus_making_2018]. In all, diplomatic visits send an important signal that the visiting and hosting countries support each other and have a close relationship.

Visiting leaders also can increase awareness of themselves among citizens in the host country. @goldsmith_spinning_2009 find that the impact of high-level government visits on foreign publics' opinions is contextual. US visits had large and positive impacts until the international media reported the negative aspects of the "war on terror" in Iraq. Again, credibility of a country's leadership is the key to conditioning the effects of high-level visits on the foreign public's perception of the country's policy. The most recent finding shows that high-level diplomatic visits across the board can improve the public approval of the visiting leader's job performance by a significantly large effect. The positive messages are especially contingent on how the public-diplomacy activities are mentioned in the news in the host countries [@goldsmith_does_2021].

# Hypotheses

We have made the case that Beijing's international development finance serves, among a diverse set of objectives, the promotion of diplomatic ties abroad and a particular image of China and its role in the international community. But, we have also highlighted a puzzle. If policymakers in China seek to support these goals with foreign aid, why the lack of transparency about where and how much China spends on development finance? We contend that limited transparency is much less a purposeful choice but a product of a complex bureaucracy that makes data reporting logistically impractical. Nonetheless, if China's foreign aid does indeed support this pair of objectives we should expect its giving to parallel other activities that are visible acts of diplomacy and image building.

In this paper, we focus on the special roles that state-sponsored media outreach and diplomatic visits have in promoting international legitimacy and stronger diplomatic ties with "southern" country governments. In addition to being visible actions taken by Beijing to support these goals, envoys and media coverage are also costly due to limited resources in terms of time and funding. We expect, therefore, that Beijing will choose on whom to cast the spotlight, and whom to visit, strategically. In particular, countries that receive higher state-sponsored media coverage and that host a greater number of diplomatic visits should be strategically important in the eyes of Beijing.

However, while public diplomacy is one way to increase a state's soft power, in the end, visits and media are fleeting and symbolic. To reinforce the objectives pursued through these visible acts of diplomacy, we argue that Beijing uses other tools, such as foreign aid, to provide material support to important countries. Hence, we expect that to the extent that Beijing uses foreign aid as a tool in support of these particular foreign policy goals, we should see Chinese development finance parallel these different but complementary activities. This rationale leads to our hypotheses:

> *H1: When developing countries receive more media exposure in* Xinhua*, they are more likely to receive more foreign aid from China.*

> *H2: When developing countries host or visit Chinese high-level diplomatic visits more frequently, they are more likely to receive more foreign aid from China.*

# Data and Methods

To assess complementarities between China's bilateral diplomatic visits and state sponsored media coverage of developing countries with Beijing's bilateral development finance, we combine data from multiple sources. We merge AidData's Chinese development finance dataset [@dreherEtAl2021 and @dreherEtAl2022] with two newly available datasets: (1) AidData's compiled yearly counts of bilateral official diplomatic visits from China [@aidData2021] and (2) millions of English edition *Xinhua* news articles from 2000 to 2014 scraped by the Cline Center for Advanced Social Research [-@cline2022]. The full dataset consists of unique recipient-year observations from 2000 to 2014.[^7]

[^7]: Version 1.0 of the Chinese development finance dataset may be obtained at: <http://aiddata.org/data/chinese-global-official-finance-dataset>. The data on China's public diplomacy can be obtained at: <http://china-dashboard.aiddata.org/>. Finally, the Archer portal used to access extracted features in the Global News Index at the Cline Center can be found at: <https://databank.illinois.edu/datasets/IDB-2179938>.

To normalize bilateral aid data, counts of country mentions, and the number of diplomatic visits, we apply the inverse hyperbolic sine transformation (ihs). We opt for this rather than the natural log because ihs is defined at zero (there are many zero observations for each of these variables) but is close enough in form to the natural log that estimates may still be interpreted as eslasticities or percent changes.

We include several controls in analysis to adjust for sources of confounding and also to improve the precision of our estimates. Prior research points to a number of covariates that are relevant.

The first set of controls consist of aid recipient characteristics. From the Penn World Table (version 9.1), we use real gross domestic product (GDP) in millions of 2011 USD [@feenstraEtAl2015].[^8] This serves as a proxy for the size of a recipient's economy. From the Penn World Table we also use population in millions and the number of people employed in millions. We normalize values using ihs to keep the transformation of the data consistent with that of the response. From the Institute for Health Metrics and Evaluation we use a yearly count of the number of individuals killed by natural disasters, and we again apply the ihs transformation [-@gbdcn2021].[^9] From PRIO/UCDP we use a binary measure of the presence of civil war, coded as 1 for each year where there is armed violence between two or more parties where one of the parties is the government and at least 25 battle-related deaths occur [@gleditschEtAl2002; @petterssonEtAl2021].[^10] To capture political and civil liberties we use Freedom House and adjust values so that higher values denote greater levels of freedom.[^11] We use this measure instead of Polity to avoid collinearities with civil war.

[^8]: Available at <https://doi.org/10.15141/S50T0R>.

[^9]: Available at <https://ghdx.healthdata.org/gbd-results-tool>.

[^10]: Available at <https://ucdp.uu.se/downloads/index.html#armedconflict>.

[^11]: Available at <https://freedomhouse.org/>.

We further control for bilateral characteristics between China and aid recipients. We account for bilateral distance in kilometers between China and developing countries using the distance measure from CEPII's gravity dataset [@mayerZignago2011]. To adjust for economic interests, we account for bilateral trade, data for which comes from CEPII's TRADEHIST dataset [@fouquinHugot2016].[^12] The former is log-transformed, while the latter is transformed via the inverse hyperbolic sine. Finally, to account for geostrategic interests, we use the alliance measure from the ATOP database, which takes the value 1 if Beijing and a recipient share an formal military alliance [@leedsEtAl2002].[^13]

[^12]: Both available at <http://www.cepii.fr/cepii/en/bdd_modele/bdd.asp>.

[^13]: Available at <http://www.atopdata.org/>.

For our empirical strategy, we estimate models of the form $$\text{ihs(aid}_{rt}) = \tau_t + \beta_1\text{ihs(coverage}_{rt}) + \beta_2\text{ihs(visits}_{rt}) + X_{rt}^\top \gamma + \epsilon_{rt}. \tag{1}$$ In the above, subscript $r$ indexes the aid recipient and $t$ the year. The vector of parameters $\tau_t$ denotes fixed year intercepts, the vector $X_{rt}$ is the set of recipient and bilateral controls, and $\epsilon_{rt}$ is a zero-centered stochastic term.

We estimate the parameters of equation 1 using a variety of techniques to ensure the robustness of our estimates. First, because coverage and diplomatic visits may be endogenous to aid, we lag each by one year (along with all time-varying covariates). We then estimate the model via OLS and report robust standard errors clustered by recipient.

However, because there are a significant number of zero observations in the outcome, OLS may yield downward biased estimates of the parameters of interest. We thus additionally estimate equation 1 using the Tobit maximum likelihood estimator. To account for unobserved heterogeneity between, and dependence within, recipients we include random recipient intercepts in the model.

In the event that the process that determines whether countries receive aid from China differs from the process that determines the amount of aid given, we further separate the estimation into selection and level equations as outlined by @cragg1971. The selection equation is estimated using mixed effects logit with random recipient intercepts and the level equation is estimated using OLS with cluster-robust standard errors.

Finally, because within-year coverage and visits may have greatest relevance for predicting the response, using the lag of each variable in model estimation may under-estimate the complementarities between coverage and visits with aid. However, estimating the models using same-year coverage and visits poses a challenge due to likely endogeneity between them and the response. Unable to find a satisfactory set of instruments, we take two different approaches. For one, we leverage the variation in the one year lag of each predictor to instrument for in-year values of each. This approach has been applied elsewhere in the literature when dealing with time-series data [see @braumoeller2008]. For the second approach, we use the internal instruments approach outlined by @lewbel2012, which leverages heteroskedasticity in the data to synthesize instrumental variables.

We present results from each of these approaches in the following analysis.

# Results

```{r}
library(tidyverse)
library(kableExtra)
dt <- read_csv(
  paste0(str_remove(getwd(), "/04_paper"), 
         "/01_data/final_data/final_data_imputed.csv")
)
```

Results are shown for `r scales::comma(nrow(dt %>% filter(year!=2000)))` unique recipient-year observations from 2001 to 2014. Multiple imputation for missing values in covariates was done via chaining random forests combined with predictive mean matching, which ensures imputed values do not differ from values already present in the data (such as in the case of a binary variable) and that the conditional distributions have realistic variances.[^14] Summary statistics are shown the Appendix.

[^14]: For multiple imputation we used the `missRanger` R package. See @stekhovenBuhlmann2011 for more on the method and its robustness relative to other approaches.

```{r, include=FALSE}
load(
  paste0(stringr::str_remove(getwd(), "/04_paper"), "/03_analysis/tab1.R")
)
load(
  paste0(stringr::str_remove(getwd(), "/04_paper"), "/03_analysis/model_list1.R") 
)
reg.tab1 <- reg.tab
cat(reg.tab1)
```

```{=tex}
\begin{table}
\caption{Model Estimates}
\begin{center}
\begin{tabular}{l c c c c c c}
\hline
 & OLS & ML Tobit & Selection & Level & IV Lag & IV Lewbel \\
\hline
Coverage (lag)              & $0.63^{*}$      & $1.29^{**}$    & $0.18^{***}$  & $0.11$          &                 &                 \\
                            & $(0.28)$        & $(0.43)$       & $(0.05)$      & $(0.14)$        &                 &                 \\
Coverage (in-year)          &                 &                &               &                 & $0.67^{*}$      & $0.51^{\cdot}$  \\
                            &                 &                &               &                 & $(0.31)$        & $(0.29)$        \\
Diplomatic Visits (lag)     & $0.93^{**}$     & $1.54^{\cdot}$ & $0.21^{**}$   & $0.36^{*}$      &                 &                 \\
                            & $(0.34)$        & $(0.83)$       & $(0.07)$      & $(0.15)$        &                 &                 \\
Diplomatic Visits (in-year) &                 &                &               &                 & $1.08^{**}$     & $1.38^{**}$     \\
                            &                 &                &               &                 & $(0.41)$        & $(0.45)$        \\
Freedom House               & $-0.36$         & $-0.58$        & $-0.08^{*}$   & $-0.20^{*}$     & $-0.36$         & $-0.39$         \\
                            & $(0.28)$        & $(0.40)$       & $(0.04)$      & $(0.09)$        & $(0.28)$        & $(0.29)$        \\
GDP                         & $-2.71^{***}$   & $-4.83^{***}$  & $-0.77^{***}$ & $0.01$          & $-2.73^{***}$   & $-2.68^{***}$   \\
                            & $(0.36)$        & $(0.48)$       & $(0.06)$      & $(0.17)$        & $(0.36)$        & $(0.34)$        \\
Population                  & $2.52^{***}$    & $5.26^{***}$   & $0.69^{***}$  & $0.27$          & $2.51^{***}$    & $2.56^{***}$    \\
                            & $(0.35)$        & $(0.76)$       & $(0.06)$      & $(0.17)$        & $(0.35)$        & $(0.36)$        \\
Unemployment                & $0.18$          & $0.84$         & $0.13$        & $-0.05$         & $0.19$          & $0.20$          \\
                            & $(0.47)$        & $(0.95)$       & $(0.08)$      & $(0.19)$        & $(0.47)$        & $(0.47)$        \\
Disaster Deaths             & $0.18$          & $0.14$         & $0.04$        & $0.02$          & $0.18$          & $0.17$          \\
                            & $(0.17)$        & $(0.27)$       & $(0.03)$      & $(0.05)$        & $(0.17)$        & $(0.16)$        \\
Civil War                   & $-1.62^{\cdot}$ & $-2.77^{*}$    & $-0.38^{*}$   & $-0.75^{\cdot}$ & $-1.62^{\cdot}$ & $-1.57^{\cdot}$ \\
                            & $(0.91)$        & $(1.25)$       & $(0.15)$      & $(0.39)$        & $(0.92)$        & $(0.91)$        \\
Trade                       & $0.14$          & $-0.18$        & $0.04$        & $0.03$          & $0.14$          & $0.14$          \\
                            & $(0.15)$        & $(0.18)$       & $(0.03)$      & $(0.05)$        & $(0.15)$        & $(0.15)$        \\
Alliance                    & $0.36$          & $2.00$         & $0.29$        & $-0.64$         & $0.07$          & $-0.32$         \\
                            & $(0.94)$        & $(2.75)$       & $(0.22)$      & $(0.47)$        & $(0.94)$        & $(1.00)$        \\
Distance                    & $1.84^{*}$      & $4.12^{\cdot}$ & $0.49^{***}$  & $0.55$          & $1.94^{*}$      & $2.21^{*}$      \\
                            & $(0.83)$        & $(2.34)$       & $(0.15)$      & $(0.37)$        & $(0.87)$        & $(0.87)$        \\
\hline
Adj. R$^2$                  & $0.19$          &              &             & $0.11$          & $0.20$          & $0.20$          \\
Num. obs.                   & $1932$          & $1932$         & $1932$        & $982$           & $1932$          & $1932$          \\
Deviance explained          &               &              & $0.15$        &               &               &               \\
\hline
\multicolumn{7}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$; $^{\cdot}p<0.1$}}
\end{tabular}
\label{table:coefficients}
\end{center}
\end{table}
```
Table 1 reports the regression coefficients for equation 1 using each of the estimation techniques outlined in the previous section. Estimates for the controls comport with prior research, save that the coefficients for democracy using the Freedom House measure reach conventional levels of statistical significance in the selection and level estimates. Contra @dreherEtAl2018, these estimates suggest China disproportionately targets its development finance to authoritarian regimes. While this result is not statistically significant for the alternative estimators, the coefficient remains consistently negative across all estimation techniques.

```{r eval=FALSE, include=FALSE}
coef_table <- map_dfr(
  model_list,
  ~ estimatr::tidy(lmtest::coeftest(.x))
) %>%
  filter(
    str_detect(term, "count") | 
      str_detect(term, "visits")
  ) %>%
  mutate(
    conf.low = estimate - 1.96 * std.error,
    conf.high = estimate + 1.96 * std.error,
    term = rep(c("Coverage", "Visits"), len = n())
  ) %>%
  select(term, estimate, conf.low, conf.high)
tibble(
  "1" = coef_table$term[1:2],
  "2" = coef_table$estimate[1:2],
  "3" = " ",
  "4" = coef_table$estimate[3:4],
  "5" = " ",
  "6" = coef_table$estimate[5:6],
  "7" = " ",
  "8" = coef_table$estimate[7:8],
  "9" = " ",
  "10" = coef_table$estimate[9:10],
  "11" = " ",
  "12" = coef_table$estimate[11:12],
  "13" = " "
) %>% 
  mutate_if(is.numeric, function(x) round(x, 2)) %>%
  kable(booktabs = T,
        format = "latex",
        caption = "Table 1: Model Estimates",
        col.names = c(
          " ", rep(c("Est.", "95% CI"), len = 12)
        )) %>%
  kable_classic(full_width = F) %>%
  column_spec(
    3, image = spec_pointrange(
      x = coef_table$estimate[1:2],
      xmin = coef_table$conf.low[1:2],
      xmax = coef_table$conf.high[1:2],
      vline = 0
    )
  ) %>%
  column_spec(
    5, image = spec_pointrange(
      x = coef_table$estimate[3:4],
      xmin = coef_table$conf.low[3:4],
      xmax = coef_table$conf.high[3:4],
      vline = 0
    )
  ) %>%
  column_spec(
    7, image = spec_pointrange(
      x = coef_table$estimate[5:6],
      xmin = coef_table$conf.low[5:6],
      xmax = coef_table$conf.high[5:6],
      vline = 0,
      lim = c(0, max(coef_table$conf.high[5:6]))
    )
  ) %>%
  column_spec(
    9, image = spec_pointrange(
      x = coef_table$estimate[7:8],
      xmin = coef_table$conf.low[7:8],
      xmax = coef_table$conf.high[7:8],
      vline = 0
    )
  )%>%
  column_spec(
    11, image = spec_pointrange(
      x = coef_table$estimate[9:10],
      xmin = coef_table$conf.low[9:10],
      xmax = coef_table$conf.high[9:10],
      vline = 0
    )
  ) %>%
  column_spec(
    13, image = spec_pointrange(
      x = coef_table$estimate[11:12],
      xmin = coef_table$conf.low[11:12],
      xmax = coef_table$conf.high[11:12],
      vline = 0
    )
  ) %>%
  add_header_above(
    c(
      " " = 1,
      "Linear" = 2,
      "Tobit" = 2,
      "Selection" = 2,
      "Level" = 2,
      "IV Lag" = 2,
      "IV Lewbel" = 2
    )
  )

```

Consistent with other studies, China prioritizes recipients with smaller economies and larger populations. Depending on the estimation strategy, greater distance is related to *more* aid, which is a finding that runs contrary to patterns observed for DAC donors. This result likely is driven by China's prioritization of aid giving in Sub-Saharan Africa. Finally, China seems to have a bias against giving aid in the context of ongoing conflict. Across estimation strategies, the coefficient on civil war is negative and marginally significant ($p < 0.1$) in four of the models and statistically significant ($p < 0.05$) in two.

China does not appear to discriminate on the basis of other measures of need, such as unemployment and natural disaster deaths. Further, trade and alliances, measures of material and strategic importance to Beijing, show no significant relationship with aid.

Turning to the estimates of interest, across specifications the coefficients on media coverage in *Xinhua* and the number of diplomatic visits run in the expected direction. Expressed as elasticities, the range of estimates suggest that one percent greater yearly instances of coverage in one recipient relative to another are associated with more than half a percent to over one percent greater aid received by that aid recipient. Further, one percent greater diplomatic visits is related to almost one percent to one and a half percent greater aid to that recipient relative to another. This finding is consistent with the view that China's development finance has public diplomacy and international legitimacy functions.

# Conclusion

Does Beijing intentionally keep its development finance a secret [@fuchsRudyak2019]? Or, does it use foreign aid as a public tool of diplomacy and to promote its legitimacy abroad?

In this study, we tackle this question by considering complementarities between where China gives aid with Chinese media coverage of developing countries intended for foreign audiences and China's bilateral diplomatic activities. These measures capture two distinct but related aspects of China's foreign policy: status/legitimacy in the eyes of a foreign (predominantly Western) audience and south-south diplomacy. To the extent that China's development finance complements these goals, we expect greater coverage of a developing country in a Chinese media outlet directed to foreign readers and greater bilateral diplomatic activity to correlate with greater Chinese foreign aid giving. To test this expectation we merge AidData's Chinese development finance dataset with two newly available datasets: (1) AidData's compiled yearly counts of bilateral official diplomatic visits from China and (2) millions of English edition *Xinhua* news articles from 2000 to 2014 scraped by the Cline Center for Advanced Social Research. We find that greater media coverage of developing countries targeted at a foreign audience via the English version of *Xinhua* and the number of diplomatic visits to a developing country predict greater receipt of Chinese aid. Our results support the view that Chinese aid allocation patterns map to China's broader diplomacy and legitimacy seeking objectives on the world stage.

Much work, of course, remains to further interrogate the links between China's development finance and its strategic and diplomatic objectives. This study provides but a glimpse into the complementarities between Beijing's foreign aid and its public diplomacy. Even so, it contributes to our understanding of the strategic determinants of China's aid, which even now contains mysteries we have yet to resolve.

\clearpage

# Appendix

Table A.1 shows summary statistics for the data sample used in the main analysis.

\clearpage

```{=tex}
\captionsetup{labelformat=AppendixTables}
\setcounter{table}{0}
```
```{r}
dt %>%
  mutate(
    aid = aid / 1000000,
    count = count / 1000,
    trade = trade / 1000000000,
    gdp = gdp / 1000000,
    pop = pop / 1000000
  ) %>%
  summarize(
    across(
      c(aid, count, govt_visits, pmm_fh, gdp:unemp, disaster:dist, trade, atopally),
      list(Mean = mean, Median = median, SD = sd, Min. = min, Max. = max),
      na.rm=T
    ) 
  ) %>%
  pivot_longer(cols = everything()) %>%
  separate(
    col = name,
    into = c(" ", "stat"),
    sep = "_(?=[^_]+$)"
  ) %>%
  mutate(
    value = scales::comma(round(value, 2))
  ) %>%
  pivot_wider(
    values_from = value,
    names_from = stat
  ) %>%
  mutate(
    " " = c(
      "Aid (millions)",
      "Article Mentions (thousands)",
      "Government Visits",
      "Freedom House",
      "GDP (in 100 millions)",
      "Population (in millions)",
      "Unemployment Rate",
      "Disaster Deaths",
      "Civil War",
      "Distance (in kilometers)",
      "Trade (in millions)",
      "Allies"
    )
  ) %>%
  kbl(
    caption = "Summary Statistics (2000-2014)",
    booktabs = T,
    linesep = ""
  )
```

```{r, results='asis'}
tabs <- c(
  "ols_region_tab1.R",
  "ivlag_region_tab1.R"
)
load(
  paste0(stringr::str_remove(getwd(), "/04_paper"), 
         "/03_analysis/",tabs[1])
)
load(
  paste0(stringr::str_remove(getwd(), "/04_paper"),
         "/03_analysis/",tabs[2])
)
#cat(ols.region)
#cat(ivlag.region)
```

\clearpage

# References