### 2022-06-14 05_Presentation/aiddata_presentation.Rmd
AidData conference presentation

---
title: "Friends Get More Money, Attention, and Handshakes: Xinhua, Diplomacy, and Chinese Foreign Aid"
subtitle: "AidData Academic Roundtable"
author: ["Lucie Lu & Miles D. Williams", "*University of Illinois at Urbana-Champaign*"]
date: 'June 24, 2022'
footer: "AidData Academic Roundtable"
output: 
  powerpoint_presentation:
    reference_doc: my_template.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
```

## What We Know

```{r}
include_graphics("china_aid_logo.jpg")
```

## What We Know

China targets aid recipients on the basis of:

-   Development need (Dreher et al. 2018)

-   Who imports more Chinese goods (Dreher and Fuchs, 2015)

-   Who has more natural resources, closer voting alignment with Beijing in the UN, and more capacity to repay loans (Dreher et al., 2021)

## What We Know

-   Like Western donors, China should wield international aid, in part, to generate goodwill toward Beijing.

<!-- -->

-   We know that this has been important for the US (e.g. tsunami relief to Indonesia in 2004)

-   There is some evidence that the perception fo Beijing's influence in developing countries is growing (per differences in 2014 and 2017 waves of AidData surveys of policymakers in 126 countries)

## Puzzle

```{r}
include_graphics("low_transparency.jpeg")
```

## Puzzle

The obscurity may have more to do with capacity than intention (Dreher et al. 2018).

## Puzzle

But transparency remains.

## Research Question and Argument

**The Puzzle**: Absent traditional modes of transparency, how does Beijing leverage its international development finance in service of its diplomatic and legitimacy-seeking objectives?

## Research Question and Argument

**The Answer**: China complements other **visible** public diplomacy and legitimacy-seeking activities with greater development finance.

## Our Data

Explanatory Variables:

-   AidData's *China's Public Diplomacy Dashboard*

-   Cline Center for Advanced Social Research's *Global News Index and Extracted Features Repository*

Outcome:

-   AidData's *Global Chinese Development Finance Dataset*

## The Underlying Logic

Two dimensions of Beijing's public diplomacy:

-   "South-South" diplomatic visits

    -   Visible and symbolic show of support and aligment

-   State-sponsored *Xinhua* coverage of developing countries

    -   Agenda-setting tool and signal of Beijing's priorities for foreign audiences

## Preview

Analysis shows that:

-   More Diplomatic Visits $\to$ More Aid

-   More State-sponsored Coverage $\to$ More Aid

## Why this Matters

These results don't just support the view **that** Beijing uses aid to support diplomatic and goals and promote its agenda. They show **how** Beijing uses aid to accomplish these goals.

## Why this Matters

Visits and coverage put a spotlight on China's priorities, while aid provides material backing that these visible signals are more than hollow performances.

## Diplomacy and Media in China's Foreign Policy

**Why focus on diplomatic visits and external-facing state-run media coverage in *Xinhua*?**

## Diplomacy

-   Visiting leaders can increase the awareness of themselves and their country among citizens in the host country (Goldsmith and Horiuchi, 2009; Goldsmith et al., 2021)

-   Diplomatic visits usually come with policy agenda between two countries, and the two are expected to cooperate in various domains through their negotiations.

-   Diplomatic visit is a signal of support and closeness of countries.

## Outward Facing Media Coverage

-   What's covered in the news signals to the public what is important (McCombs and Shaw, 1972)

-   Media are critical in building states' image to international audiences (Nye, 2008)

-   Media coverage shapes public attitudes --- core objective of mediated public diplomacy. (Entman, 2008; Kiousis and Wu, 2008)

-   *Xinhua* has dual roles:

    -   Reporting news

    -   Building up China's image (e.g., Beijing's "going out" strategy --- Shambaug, 2015)

## Diplomacy and Media and China's Foreign Policy

Visits and outward media coverage are **visible** and are linked to efforts to build **strong ties** with other nations and to **shape perceptions of Beijing's priorities**.

## The Importance of Scarcity

-   Visits and coverage provide visible signals about Beijing's priorities because they are scarce commodities.

-   Visits require planning and resources to implement.

-   Writing, editing, and publishing news articles (even online) requires time and money.

-   Limited resources and bandwidth forces choices about what countries to visit and to highlight in news stories.

-   Visits and coverage reflect Beijing's **strategic priorities.**

## Here's the Bark...Where's the Bite?

-   As important as visits and coverage are for signaling Beijing's commitments and shaping global perceptions of its priorities...

-   ...they are performative acts if they are backed up by no material support.

-   Enter global development finance.

-   Beijing uses international aid to complement its visible displays of its priorities.

## Hypotheses

H1: When developing countries receive more media exposure in *Xinhua*, they receive more foreign aid from China.

H2: When developing countries host more Chinese high-level diplomatic visits, they receive more foreign aid from China.

## Data and Methods

| Variable               | Description                                                                                        | Source                                                                                               |
|------------------------|----------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------|
| Outcome                | bilateral ODA-like flows from China to developing countries                                        | AidData's *Chinese Development Finance Dataset*                                                      |
| Explanatory Variable 1 | the number of bilateral diplomatic visits hosted by recipients from Beijing                        | AidData's compiled yearly counts of bilateral official diplomatic visits from China                  |
| Explanatory Variable 2 | the number of times a developing country is mentioned in a year in the English edition of *Xinhua* | compiled from millions of *Xinhua* articles scraped by the Cline Center for Advanced Social Research |

: The Data: Unique recipient-year observations from 2001-2014

## Data and Methods

-   Use multiple regression to adjust for various controls.

-   Apply numerous estimation approaches --- multiple estimators are justifiable, but each has its flaws.

## Results

```{r}
include_graphics("results_figure.jpg")
```

## Conclusion

-   Despite limited visibility through traditional means, China's global development finance serves a very public set of objectives.

-   Beijing leverages aid in support of such goals by complementing visible acts that signal its priorities with greater foreign aid spending.

-   These findings not only support the view that China's aid is partly in service of public diplomacy...

-   ...but also show how aid supports this objective.

## Thank You

Miles D. Williams: milesdw2\@illinois.edu

Lucie Lu: lul3\@illinois.edu

## Appendix: Model and Estimation

| Item          | Description                                                                                                                                                                                    |
|---------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| Covariates    | aid recipient characteristics (GDP, presence of civil war, disasters, political and civil liberties) and bilateral characteristics between China and aid recipients (economic ties, alliances) |
| Specification | `asinh(aid) ~ asinh(coverage) + asinh(visits) + covariates + recipient + year`                                                                                                                 |
| Estimation    | OLS (CR1 errors), Tobit (random recipient intercepts), Logit/Selection (random recipient intercepts), Level/OLS (CR1 errors), Lagged IV (CR1 errors), Lewbel IV (CR1 errors).                  |