### 2022-07-31 02_clean_and_merge/Aid_articles_identified.Rmd
Add files via upload

---
title: "Aids/development news articles identified"
author: "Lucie Lu"
date: "2022-07-26"
output: pdf_document
---

```{r}

dat <- read.csv("C:/Users/Lucie Lu/Box Sync/Xinhua co-authoring project/Xinhua news articles_2000-2017/Xinhua_2016.csv")

```

```{r}
library(tidytext)
library(dplyr)
library(tm)
library(stringr)
```

```{r, eval=F}
# pre-processing
dat$title

tidy_news <- dat %>%
    select(title) %>%
    unnest_tokens("word", title)

data("stop_words")
tidy_news <- tidy_news %>%
      anti_join(stop_words)

tidy_news %>%
  count(word) %>%
    arrange(desc(n))
```

```{r}

# dictionary-based approach to subset

aids_dictionary <- c("aid", "aids", "donor",
                     "lend", "development", "investment", "growth", "support",
                         "business", "help", "construction", "education", "funds", "projects", 
                     "donate", "assist")

dat$title2 <- str_replace_all(dat$title, "[[:punct:]]", "")
sent2 <- unlist(strsplit(dat$title2, "\\."))
aids_news_title <- grep(paste0('\\b', aids_dictionary, '\\b', collapse = '|'), sent2, ignore.case = TRUE, value = TRUE)
pattern <- paste0(aids_news_title, collapse = "|")

```

```{r}
#check how long a code runs...
library(tictoc)
tic("processing")
aids_news3 <- dat[str_detect(dat$title2, pattern),]
toc()

#aids_news <- dat[str_detect(dat$title[1:100], paste(aids_news_title, collapse="|")),]
#sent[grepl(pattern, sent)]
#later on refine the search as extracted_organizations including China or Chinese

```

```{r}

#write.csv(aids_news3, "aids_news_2016.csv")

```