### 2022-07-31 02_clean_and_merge/Pre-process China related and country counts.Rmd

---
title: "Pre-process China related and country counts"
author: "Lucie Lu"
date: "2022-07-29"
output: pdf_document
---

```{r}
#list.files("Xinhua_news_articles_2000-2014/")
df2 <- read.csv("C:/Users/Lucie Lu/Box Sync/Xinhua co-authoring project/Xinhua news articles_economy_related/Xinhua_2000_2017_aids_related.csv")

library(tidyverse)
#https://hbctraining.github.io/Intro-to-R/lessons/08_intro_tidyverse.html
df2  <- df2 %>% 
  rownames_to_column() %>% 
  as_tibble()

```

```{r}

#replace empty country with extracted locations
#df2$country <- ifelse(df2$country==" ", df2$extracted_locations, df2$country)

country2 <- matrix(NA, nrow=nrow(df2), ncol=1)
for (i in 1:nrow(df2)){
  country2[i,] <- ifelse(df2$country[i]=="", df2$extracted_locations[i], df2$country[i])
}

country2 <- as.data.frame(country2)
df2 <- cbind(df2, country2)
colnames(df2)[14] <- "country2_completed" 
```

```{r}
# identify China-related articles
#df2$country2_completed
China_dictionary <- c("People's Republic of China", "China", "Shanghai", "Beijing", "province")

df2$China_related <- ifelse(grepl(paste0('\\b', China_dictionary, '\\b', collapse = '|'), df2$country2_completed, China_dictionary)==T, 1, 0)
table(df2$China_related) #38618 related articles

```
```{r create years}

df2$year <- case_when(
  df2$column_label == 1 ~ "2000",
  df2$column_label == 2 ~ "2001",
  df2$column_label == 3 ~ "2002",
  df2$column_label == 4 ~ "2003",
  df2$column_label == 5 ~ "2004",
  df2$column_label == 6 ~ "2005",
  df2$column_label == 7 ~ "2006",
  df2$column_label == 8 ~ "2007",
  df2$column_label == 9 ~ "2008",
  df2$column_label == 10 ~ "2009",
  df2$column_label == 11 ~ "2010",
  df2$column_label == 12 ~ "2011",
  df2$column_label == 13 ~ "2012",
  df2$column_label == 14 ~ "2013",
  df2$column_label == 15 ~ "2014",
  df2$column_label == 16 ~ "2015",
  df2$column_label == 17 ~ "2016",
  df2$column_label == 18 ~ "2017",

)

table(df2$column_label)
table(df2$year)
df2$year <- as.factor(df2$year)
```

```{r}
df2_subset_China <- subset(df2, df2$China_related==1) #38,618
```

```{r, eval=F}

#write.csv(df2, "Xinhua_2000_2017_aids_news_processed.csv")
#write.csv(df2_subset_China, "Xinhua_2000_2017_aids_news_China_subset_processed.csv")
```

##########################################

```{r import multiple .csv files, eval=F}
#https://stackoverflow.com/questions/11433432/how-to-import-multiple-csv-files-at-once

temp = list.files(pattern="*.csv")
for (i in 1:length(temp)) assign(temp[i], read.csv(temp[i]))
```

```{r}
Xinhua_2000 <- subset(df2_subset_China, df2_subset_China$year==2000)
#head(Xinhua_2000)
library(tidyverse)
#https://hbctraining.github.io/Intro-to-R/lessons/08_intro_tidyverse.html
#Xinhua_2000 <- Xinhua_2000 %>% 
#  rownames_to_column() %>% 
#  as_tibble()
```

```{r}
df <- Xinhua_2000  %>% 
  pull(country2_completed) %>% 
  str_split("[|]")
#max(lengths(df))
names(df) <- Xinhua_2000$rowname
#df_2 <- data.frame(matrix(unlist(df), nrow=max(lengths(df)), byrow=TRUE),stringsAsFactors=FALSE)
library (plyr)
df_2 <- ldply (df, data.frame)
df_2 <- tibble(df_2)
colnames(df_2)[2] <- "country_mentioned"
df_2$countries_mentioned <- as.factor(df_2$country_mentioned)
```

```{r}
library(dplyr)
new_df_2000 <- df_2 %>% 
  group_by(country_mentioned) %>%
  dplyr::summarise(n = n()) %>%
#  df_2[!(df_2$country_mentioned == "")]  %>%
#  mutate(freq = round(n/sum(n), 3)) %>%
  arrange(desc(n))

new_df_2000 <- new_df_2000 %>% mutate(year==2000)

new_df_2000$year <- 2000

#duplicated name summarise in both packages
#https://statisticsglobe.com/r-error-must-only-be-used-inside-dplyr-verbs
```

```{r get frequency table}

new_df_2000 <- new_df_2000%>% 
  group_by(country_mentioned) %>%
#  dplyr::summarise(n2 = n()) %>%
  mutate(freq_perc = round(100 * n / nrow(Xinhua_2000), digits=2)) #total number of reports in 2000

```

```{r}
# list all of the files here
new_df_2000
new_df_2001
new_df_2002
new_df_2003
new_df_2004
new_df_2005
new_df_2006
new_df_2007
new_df_2008 
new_df_2009
new_df_2010
new_df_2011
new_df_2012
new_df_2013
new_df_2014
new_df_2015
new_df_2016
new_df_2017
```

```{r}
all_df <- rbind(new_df_2000,
new_df_2001,
new_df_2002,
new_df_2003,
new_df_2004,
new_df_2005,
new_df_2006,
new_df_2007,
new_df_2008,
new_df_2009,
new_df_2010,
new_df_2011,
new_df_2012,
new_df_2013,
new_df_2014,
new_df_2015,
new_df_2016,
new_df_2017)
str(all_df)
```

```{r}
#write.csv(all_df, "total.csv")

```

```{r}
#transform to wide data for better graphing
#https://cran.r-project.org/web/packages/data.table/vignettes/datatable-reshape.html
library(data.table)
county_counts_long <- reshape2::dcast(all_df, country_mentioned ~ year, value.var = "n")
head(data_all_df_long, 50)

county_freq_long <- reshape2::dcast(all_df, country_mentioned ~ year, value.var = "freq_perc")
head(county_freq_long, 50)
```

```{r}

#install.packages("countrycode")
library(countrycode)
codelist$country.name.de
```

```{r export data}
#write.csv(county_counts_long, "Xinhua_2001_2017_aids_related_articles_countrycounts.csv")
```

```{r export data frequency}
#write.csv(county_freq_long, "Xinhua_2001_2017_aids_related_articles_countryfrequencies.csv")
```