### 2022-07-31 02_clean_and_merge/combine files.Rmd

---
title: "Combine files"
author: "Lucie Lu"
date: "2022-07-28"
output: pdf_document
---

```{r setup}
# problem setting up working directory
# https://stackoverflow.com/questions/44816945/why-setting-working-directory-in-r-notebook-doesnt-change-working-directory-in
require("knitr")
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/Lucie Lu/Box Sync/Xinhua co-authoring project/Xinhua news articles_economy_related")) 
getwd()
```

```{r}
temp <- list.files(pattern="*.csv")
listOfDataFrames <- lapply(temp, read.csv)
```

```{r}

df2 <- bind_rows(listOfDataFrames, .id = "column_label")
#115,675 observations

```

```{r}

#write.csv(df2, "Xinhua_2000_2017_aids_related.csv")

```